---
title: "p8105_hw6_jmm2511"
author: "Jessie Merrifield"
date: "2023-11-30"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(readxl)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 7,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis")

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 2

Using Central Park weather data

```{r q2 data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```


```{r lm estimates and bootstrap}
q2_estimates = weather_df %>% 
  modelr::bootstrap(10, id = "strap_number") %>% 
  mutate(
    models = map(.x = strap, ~lm(tmax ~ tmin + prcp, data=.x)),
    tidy_results = map(models, broom::tidy),
    glance_results = map(models, broom::glance)
  ) %>% 
  select(strap_number, tidy_results, glance_results) %>% 
  unnest(tidy_results, glance_results) %>% 
  group_by(strap_number) %>% 
  summarise(log_product = log(estimate[term == "tmin"] * estimate[term == "prcp"]),
            r_squared=r.squared)%>%
  filter(!is.nan(log_product), !is.nan(r_squared))
#removed NaN values
    
q2_estimates  
  
```


a simple linear regression with tmax as the response with tmin and prcp as the predictors, and are interested in the distribution of two quantities estimated from these data:

r^2
log(β^1∗β^2)

Use 5000 bootstrap samples and, for each bootstrap sample, produce estimates of these two quantities. 


Plot the distribution of your estimates, and describe these in words. 

```{r plot r2}
q2_plot_r2 <- q2_estimates %>% 
  ggplot(aes(x=r_squared))+
  geom_density()

q2_plot_r2

```



```{r plot log}
q2_plot_r2 <- q2_estimates %>% 
  ggplot(aes(x=log_product))+
  geom_density()

q2_plot_r2

```


Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for r^2
 and log(β^0∗β^1)

```{r CI}
#bootstrap_results |> 
 # group_by(term) |> 
  #summarize(
   # ci_lower = quantile(estimate, 0.025), 
    #ci_upper = quantile(estimate, 0.975))
```



. Note: broom::glance() is helpful for extracting r^2
 from a fitted regression, and broom::tidy() (with some additional wrangling) should help in computing log(β^1∗β^2)


## Problem 3

In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. 

Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).

```{r q3 read in data}
birthweight<-read.csv("./data/birthweight.csv")
```

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.



Compare your model to two others:



One using length at birth and gestational age as predictors (main effects only)



One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
Make this comparison in terms of the cross-validated prediction error; use crossv_mc and functions in purrr as appropriate.

